MS Keyrun Species, Environment, Data Formats

### 2020-07-31 docs/SpeciesEnvDat.Rmd
update with proposed decsions and date

Current species groups in the GB Rpath model:
```{r rpathspp, message=FALSE}

@@

source_data("https://github.com/NOAA-EDAB/GBRpath/blob/master/data/SOE_species_list.RData?raw=True")

@@

@@

@@

How many models are implemented in R? See [discussion](https://docs.google.com/document/d/19FpJ8w2a-ZZ8ZY4ZmsnjlIilLeN1X9_Tz1dM-KsdOOw/edit#heading=h.kvnvc0pxyvqq) of pros and cons of using R data package for standardizing input datasets and making them available.

@@

### Survey conducted July 13-24 2020

@@

```{r getsurvey}

@@

Responses:
```{r, fig.height=6}
# each row is a response with a comma delimited list of species
# string split each row to species
# tally by species
# histogram

@@

Responses:
```{r, fig.height=10}
# each row is a response with a comma delimited list of environmental data
# string split each row to dataset name
# tally by dataset
# histogram

@@

Responses:
```{r}
# these are multiple choice yes/no so just plot

useR <- data.frame(ModelInR = responses$ModelInR,
                   ModelUseR = responses$ModelUseR,
                   RData = responses$RData) %>%
  gather(question, answer, ModelInR:RData) %>%
  mutate(shortans = ifelse(is.na(word(answer, 1, 2)), 
                           answer, 
                           word(answer, 1, 2)))

@@

```
Short answer key:
"No but" = `r unique(useR$answer)[2]`  
"Probably, we" = `r unique(useR$answer)[4]`

>>>>

(We first pull species lists from the Rpath model and load the survey results.)

```{r getrpathspp, message=FALSE}

@@

invisible(source_data("https://github.com/NOAA-EDAB/GBRpath/blob/master/data/SOE_species_list.RData?raw=True"))

@@

```{r getsurvey, message=FALSE}

These species received >1 vote for inclusion in MS-Keyrun multispecies models:

```{r msmodspp}
# each row is a response with a comma delimited list of species
# string split each row to species
# tally by species
# histogram

focal <- msspecies %>%
  filter(Model == "MS") %>%
  count(Species) %>%
  filter(n > 1) %>%
  left_join(rpathspp, by = c("Species" = "RPATH")) %>%
  select(-COMNAME) %>%
  arrange(desc(n))

knitr::kable(focal)

Species modeled by all compared multispecies models could include all of these, or could be limited to a subset.

Current species groups in the GB Rpath model (all retained for MS-Keyrun):
```{r rpathspp, message=FALSE}

@@

Hydra, the length-based multispecies model, already incorporates a time series of Georges Bank bottom temperature as a covariate in simulation mode to force changes in growth, maturity, and recruitment. For MS-Keyrun, we can explore fitting the parameters governing the influence of this covariate.

Rpath can be directly forced by a primary production time series, and temperature-based mediation functions can be included (but it is unclear whether parameters can be fit for these functions).

Two mulispecies models are not currently set up to incorporate environmental data:
MSCAA, MSSPM/Kraken

The WHAM model can potentially incorporate any environmental time series that can influence recruitment or mortality.

These environmental time series received >1 vote for inclusion in MS-Keyrun multispecies and Rpath models:
```{r envdat}
# each row is a response with a comma delimited list of environmental data
# string split each row to dataset name
# tally by dataset
# histogram

includedenv <- msenv %>%
  count(Environment) %>%
  filter(n > 1) %>%
  arrange(desc(n))

knitr::kable(includedenv)

We can include GB temperature and primary production time series as part of the MS-Keyrun dataset. Modelers may retrieve additional data from `ecodata`.

@@

Most models are either implemented in R or can work directly with R datasets:
```{r useRdata}
# these are multiple choice yes/no so just plot

useR <- data.frame(ModelInR = responses$ModelInR,
                   ModelUseR = responses$ModelUseR,
                   RData = responses$RData) %>%
  gather(question, answer, ModelInR:RData) %>%
  mutate(shortans = ifelse(is.na(word(answer, 1, 2)), 
                           answer, 
                           word(answer, 1, 2)))
yesR <- useR %>%
  group_by(question) %>%
  count(answer=="Yes") %>%
  mutate(tot = sum(n)) %>%
  mutate(percent = n/tot*100)

knitr::kable(yesR)

All surveyed agreed that we should maintain datasets within an R data package for the MS-Keyrun project.

See [discussion](https://docs.google.com/document/d/19FpJ8w2a-ZZ8ZY4ZmsnjlIilLeN1X9_Tz1dM-KsdOOw/edit#heading=h.kvnvc0pxyvqq) of pros and cons of using R data package for standardizing input datasets and making them available.

@@

### Survey conducted July 13-29 2020

@@

@@

Responses:
```{r plotspp, fig.height=6}

@@

Responses:
```{r plotenv, fig.height=10}

@@

Responses:
```{r plotdat}

@@

Short answer key:
"No but" = `r unique(useR$answer)[2]`  
"Probably, we" = `r unique(useR$answer)[4]`
